---
title: "Journal Tone"
output: 
  learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(scales)
library(topicmodels)
library(scifetch)
library(tidytext)
knitr::opts_chunk$set(echo = FALSE)
```


## Journal based

### Usage

Change the journal ISSN number before [TA] and the time range before[DP] to get the records from pubmed. I limited the numbers of records to 10000. Since such fetch from Pubmed might be blocked, try to use it carefully. I suggest every query should be limited no more than 8000, which is enough to show some trends.

Just click `Run Code` to get the results

```{r journal, exercise = T, exercise.timelimit = 500}
query <- '0003-2700[TA] AND 2017/06:2017/08[DP]'
# fetch data
tmdf <- getpubmed(query, start = 1, end = 10000) %>%
        getpubmedtbl() %>%
        mutate(time = as.POSIXct(date, origin = "1970-01-01"),
         month = round_date(date, "month"))
# get words in title
wordft <- tmdf %>%
        filter(nchar(title) > 0) %>%
        unnest_tokens(word, title,drop = F) %>%
        anti_join(stop_words) %>%
        filter(str_detect(word, "[^\\d]")) %>%
        group_by(word) %>%
        mutate(word_total = n()) %>%
        ungroup() %>%
        mutate(source = 'title')
# plot top 20 words in title
wordft %>%
        count(word, sort = TRUE) %>%
        top_n(20,n) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
  geom_col(show.legend = FALSE) +
  ylab("Top 20 commonly used words in titles") +
  coord_flip()
# get words in abstracts
wordfabs <- tmdf %>%
        filter(nchar(abstract) > 0) %>%
        unnest_tokens(word, abstract,drop = F) %>%
        anti_join(stop_words) %>%
        filter(str_detect(word, "[^\\d]")) %>%
        group_by(word) %>%
        mutate(word_total = n()) %>%
        ungroup() %>%
        mutate(source = 'abstract')
        
# plot top 20 words in abstracts
wordfabs %>%
        count(word, sort = TRUE) %>%
        top_n(20,n) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
  geom_col(show.legend = FALSE) +
  ylab("Top 20 commonly used words in abstracts") +
  coord_flip()
        
        
# Temporal Trends
papers_per_month <- tmdf %>%
  group_by(month) %>%
  summarize(month_total = n())
# Growing words in titles
word_month_counts <- wordft %>%
  filter(word_total >= 100) %>%
  count(word, month) %>%
  complete(word, month, fill = list(n = 0)) %>%
  inner_join(papers_per_month, by = "month") %>%
  mutate(percent = n / month_total) %>%
  mutate(year = year(month) + yday(month) / 365) %>%
        filter(percent < 0.8)

mod <- ~ glm(cbind(n, month_total - n) ~ year, ., family = "binomial")

slopes <- word_month_counts %>%
  nest(-word) %>%
  mutate(model = map(data, mod)) %>%
  unnest(map(model, tidy)) %>%
  filter(term == "year") %>%
  arrange(desc(estimate))

slopes %>%
  head(9) %>%
  inner_join(word_month_counts, by = "word") %>%
  mutate(word = reorder(word, -estimate)) %>%
  ggplot(aes(month, n / month_total, color = word)) +
  geom_line(show.legend = FALSE) +
  scale_y_continuous(labels = percent_format()) +
  facet_wrap(~ word, scales = "free_y") +
  expand_limits(y = 0) +
  labs(x = "Year",
       y = "Percentage of titles containing this term",
       title = "9 fastest growing words",
       subtitle = "Judged by growth rate"
              )
# shrinking words in titles
slopes %>%
  tail(9) %>%
  inner_join(word_month_counts, by = "word") %>%
  mutate(word = reorder(word, estimate)) %>%
  ggplot(aes(month, n / month_total, color = word)) +
  geom_line(show.legend = FALSE) +
  scale_y_continuous(labels = percent_format()) +
  facet_wrap(~ word, scales = "free_y") +
  expand_limits(y = 0) +
  labs(x = "Year",
       y = "Percentage of titles containing this term",
       title = "9 fastest shrinking words",
       subtitle = "Judged by growth rate"
              )

# Growing words in abstracts
word_month_counts <- wordfabs %>%
  filter(word_total >= 100) %>%
  count(word, month) %>%
  complete(word, month, fill = list(n = 0)) %>%
  inner_join(papers_per_month, by = "month") %>%
  mutate(percent = n / month_total) %>%
  mutate(year = year(month) + yday(month) / 365) %>%
        filter(percent < 0.8)

mod <- ~ glm(cbind(n, month_total - n) ~ year, ., family = "binomial")

slopes <- word_month_counts %>%
  nest(-word) %>%
  mutate(model = map(data, mod)) %>%
  unnest(map(model, tidy)) %>%
  filter(term == "year") %>%
  arrange(desc(estimate))

slopes %>%
  head(9) %>%
  inner_join(word_month_counts, by = "word") %>%
  mutate(word = reorder(word, -estimate)) %>%
  ggplot(aes(month, n / month_total, color = word)) +
  geom_line(show.legend = FALSE) +
  scale_y_continuous(labels = percent_format()) +
  facet_wrap(~ word, scales = "free_y") +
  expand_limits(y = 0) +
  labs(x = "Year",
       y = "Percentage of abstracts containing this term",
       title = "9 fastest growing words",
       subtitle = "Judged by growth rate")
        
# shrinking words in abstracts
slopes %>%
  tail(9) %>%
  inner_join(word_month_counts, by = "word") %>%
  mutate(word = reorder(word, estimate)) %>%
  ggplot(aes(month, n / month_total, color = word)) +
  geom_line(show.legend = FALSE) +
  scale_y_continuous(labels = percent_format()) +
  facet_wrap(~ word, scales = "free_y") +
  expand_limits(y = 0) +
  labs(x = "Year",
       y = "Percentage of abstracts containing this term",
       title = "9 fastest shrinking words",
       subtitle = "Judged by growth rate")

# topic model
desc_dtm <- wordfabs %>%
        count(line, word, sort = TRUE) %>%
        ungroup() %>%
        cast_dtm(line, word, n)

desc_lda <- LDA(desc_dtm, k = 5, control = list(seed = 42))
tidy_lda <- tidy(desc_lda)

top_terms <- tidy_lda %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  group_by(topic, term) %>%    
  arrange(desc(beta)) %>%  
  ungroup() %>%
  mutate(term = factor(paste(term, topic, sep = "__"), 
                       levels = rev(paste(term, topic, sep = "__")))) %>%
  ggplot(aes(term, beta, fill = as.factor(topic))) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
  labs(title = "Top 10 terms in each LDA topic",
       x = NULL, y = expression(beta)) +
  facet_wrap(~ topic, ncol = 5, scales = "free")
```

## Keywords based

### Usage

Change the keywords before [MH] and the time range before[DP] to get the records from pubmed. I limited the numbers of records to 10000. Since such fetch from Pubmed might be blocked, try to use it carefully. I suggest every query should be limited no more than 8000, which is enough to show some trends.

Just click `Run Code` to get the results

```{r keywords, exercise = T, exercise.timelimit = 1000}
query <- 'Solid Phase MicroExtraction[MH] AND 1997/08:2017/08[DP]'
tmdf <- getpubmed(query, start = 1, end = 10000) %>%
        getpubmedtbl() %>%
        mutate(time = as.POSIXct(date, origin = "1970-01-01"),
         month = round_date(date, "month"))
# Numbers by journal
tmdf %>%
  group_by(journal) %>%
  summarize(papers = n_distinct(title)) %>%
  top_n(20,papers) %>%
  mutate(journal = reorder(journal, papers)) %>%     
  ggplot(aes(journal, papers)) +
  geom_col() +
  coord_flip()
# Get the words in title (journal with papers > 100)
wordft <- tmdf %>%
        filter(nchar(title) > 0) %>%
        group_by(journal) %>%
        mutate(journal_total = n()) %>%
        ungroup() %>%
        filter(journal_total > 100) %>%
        unnest_tokens(word, title,drop = F) %>%
        anti_join(stop_words) %>%
        filter(str_detect(word, "[^\\d]")) %>%
        group_by(word) %>%
        mutate(word_total = n()) %>%
        ungroup() %>%
        mutate(source = 'title')
  
# get words in abstracts
wordfabs <- tmdf %>%
        filter(nchar(abstract) > 0) %>%
        group_by(journal) %>%
        mutate(journal_total = n()) %>%
        ungroup() %>%
        filter(journal_total > 100) %>%
        unnest_tokens(word, abstract,drop = F) %>%
        anti_join(stop_words) %>%
        filter(str_detect(word, "[^\\d]")) %>%
        group_by(word) %>%
        mutate(word_total = n()) %>%
        ungroup() %>%
        mutate(source = 'abstract')
        
# plot top 20 words in abstracts
wordfabs %>%
        count(word, sort = TRUE) %>%
        top_n(20,n) %>%
        mutate(word = reorder(word, n)) %>%
        ggplot(aes(word, n)) +
  geom_col(show.legend = FALSE) +
  ylab("Top 20 commonly used words in abstracts") +
  coord_flip()
        
# Temporal Trends
papers_per_month <- tmdf %>%
  group_by(month) %>%
  summarize(month_total = n())
# Growing words in titles
word_month_counts <- wordft %>%
  filter(word_total >= 100) %>%
  count(word, month) %>%
  complete(word, month, fill = list(n = 0)) %>%
  inner_join(papers_per_month, by = "month") %>%
  mutate(percent = n / month_total) %>%
  mutate(year = year(month) + yday(month) / 365) %>%
        filter(percent < 0.8)

mod <- ~ glm(cbind(n, month_total - n) ~ year, ., family = "binomial")

slopes <- word_month_counts %>%
  nest(-word) %>%
  mutate(model = map(data, mod)) %>%
  unnest(map(model, tidy)) %>%
  filter(term == "year") %>%
  arrange(desc(estimate))

slopes %>%
  head(9) %>%
  inner_join(word_month_counts, by = "word") %>%
  mutate(word = reorder(word, -estimate)) %>%
  ggplot(aes(month, n / month_total, color = word)) +
  geom_line(show.legend = FALSE) +
  scale_y_continuous(labels = percent_format()) +
  facet_wrap(~ word, scales = "free_y") +
  expand_limits(y = 0) +
  labs(x = "Year",
       y = "Percentage of titles containing this term",
       title = "9 fastest growing words",
       subtitle = "Judged by growth rate"
              )
# shrinking words in titles
slopes %>%
  tail(9) %>%
  inner_join(word_month_counts, by = "word") %>%
  mutate(word = reorder(word, estimate)) %>%
  ggplot(aes(month, n / month_total, color = word)) +
  geom_line(show.legend = FALSE) +
  scale_y_continuous(labels = percent_format()) +
  facet_wrap(~ word, scales = "free_y") +
  expand_limits(y = 0) +
  labs(x = "Year",
       y = "Percentage of titles containing this term",
       title = "9 fastest shrinking words",
       subtitle = "Judged by growth rate"
              )

# Growing words in abstracts
word_month_counts <- wordfabs %>%
  filter(word_total >= 100) %>%
  count(word, month) %>%
  complete(word, month, fill = list(n = 0)) %>%
  inner_join(papers_per_month, by = "month") %>%
  mutate(percent = n / month_total) %>%
  mutate(year = year(month) + yday(month) / 365) %>%
        filter(percent < 0.8)

mod <- ~ glm(cbind(n, month_total - n) ~ year, ., family = "binomial")

slopes <- word_month_counts %>%
  nest(-word) %>%
  mutate(model = map(data, mod)) %>%
  unnest(map(model, tidy)) %>%
  filter(term == "year") %>%
  arrange(desc(estimate))

slopes %>%
  head(9) %>%
  inner_join(word_month_counts, by = "word") %>%
  mutate(word = reorder(word, -estimate)) %>%
  ggplot(aes(month, n / month_total, color = word)) +
  geom_line(show.legend = FALSE) +
  scale_y_continuous(labels = percent_format()) +
  facet_wrap(~ word, scales = "free_y") +
  expand_limits(y = 0) +
  labs(x = "Year",
       y = "Percentage of abstracts containing this term",
       title = "9 fastest growing words ",
       subtitle = "Judged by growth rate")
        
# shrinking words in abstracts
slopes %>%
  tail(9) %>%
  inner_join(word_month_counts, by = "word") %>%
  mutate(word = reorder(word, estimate)) %>%
  ggplot(aes(month, n / month_total, color = word)) +
  geom_line(show.legend = FALSE) +
  scale_y_continuous(labels = percent_format()) +
  facet_wrap(~ word, scales = "free_y") +
  expand_limits(y = 0) +
  labs(x = "Year",
       y = "Percentage of abstracts containing this term",
       title = "9 fastest shrinking words",
       subtitle = "Judged by growth rate")

# topic model
desc_dtm <- wordfabs %>%
        count(line, word, sort = TRUE) %>%
        ungroup() %>%
        cast_dtm(line, word, n)

desc_lda <- LDA(desc_dtm, k = 5, control = list(seed = 42))
tidy_lda <- tidy(desc_lda)

top_terms <- tidy_lda %>%
  group_by(topic) %>%
  top_n(10, beta) %>%
  ungroup() %>%
  arrange(topic, -beta)

top_terms %>%
  mutate(term = reorder(term, beta)) %>%
  group_by(topic, term) %>%    
  arrange(desc(beta)) %>%  
  ungroup() %>%
  mutate(term = factor(paste(term, topic, sep = "__"), 
                       levels = rev(paste(term, topic, sep = "__")))) %>%
  ggplot(aes(term, beta, fill = as.factor(topic))) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
  labs(title = "Top 10 terms in each LDA topic",
       x = NULL, y = expression(beta)) +
  facet_wrap(~ topic, ncol = 5, scales = "free") 

# Journal accent

words_by_journal <- wordfabs %>%
  count(journal, word, sort = TRUE) %>%
  ungroup()

tf_idf <- words_by_journal %>%
  bind_tf_idf(word, journal, n) %>%
  arrange(desc(tf_idf))

tf_idf %>%
  group_by(journal) %>%
  top_n(10, tf_idf) %>%
  ungroup() %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(word, tf_idf, fill = journal)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ journal, scales = "free") +
  ylab("tf-idf in abstracts") +
  coord_flip()
```


